Capstone Movielens Report

### 2025-12-04 reports/capstone-movielens.site/02.7.UMGYE-model.Rmd
Regularized UMGE Model description: minor fix.
Function described: `calc_date_general_effect.cv`.

#### [calc_date_general_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L2) Function {#func.calc_date_general_effect}
\

This is a helper function that calculates the _date general effect_ of the user ratings and is intended to be used in other functions and scripts (such as [train_UMGY_effect](#func.train_UMGY_effect)) described below.

The following is a slightly simplified version of the function's source code:
```{r eval=FALSE}
calc_date_general_effect <- function(train_set, lambda = 0){
  if (is.na(lambda)) {
    stop("Function: calc_date_general_effect
`lambda` is `NA`")
  }

  if(lambda == 0) put_log("Function `calc_date_general_effect`:
Computing Date Global Effect for given Train Set data...")
  else put_log1("Function `calc_date_general_effect`:
Computing Date Global Effect for lambda: %1...",
                lambda)
  dg_effect <- train_set |> 
    left_join(edx.user_effect, by = "userId") |>
    left_join(rglr.UM_effect, by = "movieId") |>
    left_join(rglr.UMG_effect, by = "movieId") |>
    left_join(date_days_map, by = "timestamp") |>
    mutate(resid = rating - (mu + a + b + g)) |>
  group_by(days) |>
    summarise(de = mean_reg(resid, lambda), 
              year = mean(year))
  dg_effect
}
```

::: {.noteblock data-latex=""}
The complete version of the source code of the [calc_date_general_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L2) function is available in the [UMGYE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L1) section of the [UMGY-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L2) script on _GitHub_.
:::

A data structure containing information about _Date General Effect_.

#### [calc_date_general_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L30) Function {#func.calc_date_general_effect.cv}
\

This is a helper function that calculates the _date general effect_ of the user ratings using the _5-Fold Cross Validation_ method and is intended to be used in other functions and scripts (such as [train_UMGY_effect.cv](#func.train_UMGY_effect.cv)) described below.

##### Souurce Code
\

::: {.noteblock data-latex=""}
The complete version of the source code of the [calc_date_general_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L30) function is available in the [UMGYE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L1) section of the [UMGY-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L30) script on _GitHub_.
:::

The following is a simplified version of the function's source code:
```{r eval=FALSE}
calc_date_general_effect.cv <- function(lambda = 0){
  if (is.na(lambda)) {
    stop("Function: calc_date_general_effect.cv
`lambda` is `NA`")
  }

  date_general_effect_ls <- lapply(edx_CV,  function(cv_fold_dat){
    cv_fold_dat$train_set |> calc_date_general_effect(lambda)
  })
  str(date_general_effect_ls)

  date_general_effect_united <- union_cv_results(date_general_effect_ls)

  date_general_effect_united |>
    group_by(days) |>
    summarise(de = mean(de, na.rm = TRUE), year = mean(year, na.rm = TRUE))
}
```

::: {.noteblock data-latex=""}
Here we use the function call [union_cv_results](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L432) to aggregate the _5-Fold Cross Validation_ results (for details, see the [union_cv_results](#func.union_cv_results) function description in the [Data Helper Functions](#appndx_a.data_helper.functions) section of [Appendix A: Support Functions]).
:::

A data structure containing information about _Date General Effect_ as result of _5-Fold Cross Validation_.

@@

Now, we have a bit better result than one for the previous model.

>>>>

@@

Now, we have a bit better result than one for the previous (*Regularized UMGE*) model.

### 2025-12-04 reports/capstone-movielens.site/05.appendix.Rmd
Function described: `calc_date_general_effect.cv`.

#### [calc_date_general_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM-effect.functions.R#L4) Function {#func.calc_date_general_effect}
\

This is a helper function that calculates the _Date General Effect_ (optionally *regularizes* if the [lambda](#func.calc_date_general_effect.args.lambda) argument value is non-default)of the user ratings and is intended to be used in other functions and scripts (such as [train_UMGY_effect](#func.train_UMGY_effect)) described below.

@@

::: {.noteblock data-latex=""}
In the sample code snippet above the function `calc_date_general_effect` is used internally in the [train_UMGY_effect](#func.train_UMGY_effect)) described below.
:::
:::

@@

Below is a simplified version of the source code for the [calc_date_general_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM-effect.functions.R#L4) function:
```{r eval=FALSE}
  dg_effect <- train_set |> 
    left_join(edx.user_effect, by = "userId") |>
    left_join(rglr.UM_effect, by = "movieId") |>
    left_join(rglr.UMG_effect, by = "movieId") |>
    left_join(date_days_map, by = "timestamp") |>
    mutate(resid = rating - (mu + a + b + g)) |>
  group_by(days) |>
    summarise(de = mean_reg(resid, lambda), 
              year = mean(year))

@@

::: {.noteblock data-latex=""}
The complete source code of the [calc_date_general_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM-effect.functions.R#L4) function is available in the [Utility Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM-effect.functions.R#L3) section of the [UM-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM-effect.functions.R#L4) script on _GitHub_.
:::
:::

#### [calc_date_general_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM-effect.functions.R#L4) Function {#func.calc_date_general_effect.cv}
\

Trains (and optionally *regularizes* if the [lambda](#func.calc_date_general_effect.cv.args.lambda) argument value is non-default) the *User+Movie Effect (UME) Model* using the training dataset passed in the [train_set](#func.calc_date_general_effect.cv.args.train_set) argument.

::: { #func.calc_date_general_effect.cv.usage .sidebar }
##### Usage
\

```{r eval=FALSE}
  UM_effect <- calc_date_general_effect.cv(train_set, lambda = 0)
```
:::

@@

The *Regularization technique* (described in section [Section 24.1 *Case study: recommendation systems / Penalized Least Squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1]) is applied to the model being trained only when the *Regularization parameter* value, passed in the [lambda](#func.calc_date_general_effect.cv.args.lambda) argument, is other than default (to be more exact, greater than zero). The regularization is committed through the *penalized regression* by computing the *penalized mean* performed by the [mean_reg](#func.mean_reg) function described above in the *Common Helper Functions /* [Regularization Functions] section of *this Appendix*.
:::

@@

Data frame representing the trained (and also *regularized* if [lambda](#func.calc_date_general_effect.cv.args.lambda) arg value other than default) *UM Effect Model*.
:::

@@

The source code of the [calc_date_general_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM-effect.functions.R#L4) function is shown below:
```{r eval=FALSE}
calc_date_general_effect.cv <- function(train_set, lambda = 0){
  if (is.na(lambda)) {
    stop("Function: calc_date_general_effect.cv
`lambda` is `NA`")
  }

::: {.noteblock data-latex=""}
The source code of the [calc_date_general_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM-effect.functions.R#L4) function is also available in the [Utility Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM-effect.functions.R#L3) section of the [UM-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM-effect.functions.R#L4) script on _GitHub_.
:::
:::

>>>>

#### [calc_date_general_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L4) Function {#func.calc_date_general_effect}
\

This is a helper function that calculates the _Date General Effect_ (optionally *regularizes* if the [lambda](#func.calc_date_general_effect.args.lambda) argument value is non-default) of the user ratings and is intended to be used in other functions and scripts (such as [train_UMGY_effect](#func.train_UMGY_effect)) described below.

@@

::: {.noteblock data-latex=""}
In the code snippet above, the function `calc_date_general_effect` is called internally from the [train_UMGY_effect](#func.train_UMGY_effect) function described below.
:::
:::

@@

Below is a slightly simplified version of the source code of the [calc_date_general_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L4) function:
```{r eval=FALSE}
calc_date_general_effect <- function(train_set, lambda = 0){
  dg_effect <- train_set |> 
    left_join(edx.user_effect, by = "userId") |>
    left_join(rglr.UM_effect, by = "movieId") |>
    left_join(rglr.UMG_effect, by = "movieId") |>
    left_join(date_days_map, by = "timestamp") |>
    mutate(resid = rating - (mu + a + b + g)) |>
  group_by(days) |>
    summarise(de = mean_reg(resid, lambda), 
              year = mean(year))

@@

::: {.noteblock data-latex=""}
The complete source code of the [calc_date_general_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L4) function is available in the [Utility Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L3) section of the [UMGY-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L4) script on _GitHub_.
:::
:::

#### [calc_date_general_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L32) Function {#func.calc_date_general_effect.cv}
\

This is a helper function that calculates the _Date General Effect_ (optionally *regularizes* if the [lambda](#func.calc_date_general_effect.cv.args.lambda) argument value is non-default) of the user ratings using the _5-Fold Cross Validation_ and is intended to be used in other functions and scripts (such as [train_UMGY_effect](#func.train_UMGY_effect)) described below.

::: { #func.calc_date_general_effect.cv.usage .sidebar }
##### Usage
\
```{r eval=FALSE}
calc_date_general_effect.cv(lambda) |> calc_UMGY_effect()  
```

::: {.noteblock data-latex=""}
In the above code snippet, the `calc_date_general_effect.cv` function is called internally from the [train_UMGY_effect.cv](#func.train_UMGY_effect.cv function) function, along with the [calc_UMGY_effect](#func.calc_UMGY_effect) function (both described below).
:::
:::

@@

Under the hood, this function internally calls the [calc_date_general_effect](#func.calc_date_general_effect) function described above for each iteration of the *K-Fold Cross-Validation* applied to the [`edx_CV` Object] described below in the [Appendix B: Models Training Datasets]. It also calls internally the [union_cv_results](#func.union_cv_results) function described above in the *Data Helper Functions / [Models Training Functions]* section of this *Appendix* to aggregate the cross-validation results.

::: {.noteblock data-latex=""}
The *Regularization technique* (described in section [Section 24.1 *Case study: recommendation systems / Penalized Least Squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1]) is applied to the *Date General Effect* calculated by this function only when the *Regularization parameter* value, passed in the [lambda](#func.calc_date_general_effect.cv.args.lambda) argument, is non-default (to be more exact, greater than zero). The regularization is committed through the *penalized regression* by computing the *penalized mean* performed by the internally called [mean_reg](#func.mean_reg) function described above in the *Common Helper Functions /* [Regularization Functions] section of *this Appendix*.
:::
:::

@@

Data frame representing (*regularized* if [lambda](#func.calc_date_general_effect.cv.args.lambda) argument value is non-default) *Date General Effect*.
:::

@@

Below is a simplified version of the source code of the [calc_date_general_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L32) function:
```{r eval=FALSE}
calc_date_general_effect.cv <- function(lambda = 0){

  date_general_effect_ls <- lapply(edx_CV,  function(cv_fold_dat){
    cv_fold_dat$train_set |> calc_date_general_effect(lambda)
  })
  put_log1("Function `calc_date_general_effect.cv`:
Date Global Effect list has been computed for %1-Fold Cross Validation samples.", 
           CVFolds_N)

  date_general_effect_united <- union_cv_results(date_general_effect_ls)

  date_general_effect <- date_general_effect_united |>
    group_by(days) |>
    summarise(de = mean(de, na.rm = TRUE), year = mean(year, na.rm = TRUE))

  if(lambda == 0) put_log("Function `calc_date_general_effect.cv`:
Training completed: Date Global Effects model.")
  else put_log1("Function `calc_date_general_effect.cv`:
Training completed: Date Global Effects model for lambda: %1...",
                lambda)

  date_general_effect
}
```

::: {.noteblock data-latex=""}
The complete source code of the [calc_date_general_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L32) function is available in the [Utility Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L3) section of the [UMGY-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L32) script on _GitHub_.
:::
:::
